---
title: "Chapter 9, 10, 11"
author: "Maddie Lombardo"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Chapter 9

# Chapter 10

## Chapter 10.5 Exercises

### Question 1: How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

```{r chapter 10.5 question 1}
print(mtcars)

as_tibble(mtcars) %>% 
  print()
```

With a tibble, the type of data in each column is expressed (i.e dbl, chr, etc). This information is not shown in a dataframe

### Question 2: Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r 10.5 Question 2}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

tbl <- tibble(abc = 1, xyz = "a")
tbl$xyz
tbl[, "xyz"]
tbl[, c("abc", "xyz")]
```

The default dataframe behaviors may be annoying because they require much less specificity and therefore could be problematic for larger datasets.

### Question 3: If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r 10.5 Question 3}
df <- tibble(var = "mpg", typ = "car")

df$var
```

### Question 4.1: Practice referring to non-syntactic names in the following data frame by: Extracting the variable called 1.
```{r 10.5 question 4.1}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

annoying[[1]]
```

### Question 4.2: Plotting a scatterplot of 1 vs 2.

```{r 10.5 question 4.2}
ggplot(data = annoying, mapping = aes(x = `1`, y = `2`)) +
  geom_point()
```

### Question 4.3: Creating a new column called 3 which is 2 divided by 1.

```{r 10.5 question 4.3}
annoying <- annoying %>% 
  mutate(`3` = `2`/`1`)
```

### Question 4.4: Renaming the columns to one, two and three.

```{r 10.5 question 4.4}
annoying <- annoying %>% 
  rename("one" = `1` , "two" = `2`,"three" = `3`)
```

### Question 5: What does tibble::enframe() do? When might you use it?

It converts named atomic vectors or lists to one- or two-column data frames with set column names. You may use this when you want to convert data in a way that you could more easily manipulate it.

### Question 6: What option controls how many additional column names are printed at the footer of a tibble?

print.tbl explains the options. There are functions, max_extra_cols,max_footer_lines, that allow you to control how many column names are at the footer.

# Chapter 11

## Chapter 11.2.2 Exercises

### Question 1: What function would you use to read a file where fields were separated with “|”?

use read_delim(file, delim = "|")

### Question 2: Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

col_names, col_types, col_select, na, id, quoted_na, name_repair

### Question 3: What are the most important arguments to read_fwf()?

file, col_positions, na, skip, name_repair

### Question 4: Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By default, read_csv() assumes that the quoting character will be ". What argument to read_csv() do you need to specify to read the following text into a data frame?

"x,y\n1,'a,b'"

you should use the escape_backslash or escape_double arguments to specify to read the text into a dataframe

### Question 5: Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r 11.2.2 Question 5.1}
read_csv("a,b\n1,2,3\n4,5,6")
```
This only creates 2 column names with 3 variables, so the 2,3 and 5,6 get smooshed together into column b.


```{r 11.2.2 question 5.2}
read_csv("a,b,c\n1,2\n1,2,3,4")
```
This one creates 3 columns: a, b, c where row 1 only has 2 values and row two has 4. This caused the data to set NA to c1 and smoosh 3,4 into c2

```{r 11.2.2 Question 5.3}
read_csv("a,b\n\"1")
```
the additional backslash sets no data to the columns a and b so no dataframe gets created

```{r 11.2.2 question 5.4}
read_csv("a,b\n1,2\na,b")
```
because there is no '\nna,b' r gets confused and sets the second row position a to be a instead of NA.

```{r 11.2.2 question 5.5}
read_csv("a;b\n1;3")
```
the semicolons have confused R studio when they should have been colons (:) further, they do not assign as they would in subsetting a tibble and will just print exactly "a:b" and "1:3" not 1-3.

## Chapter 11.3.5 Exercises

### Question 1: What are the most important arguments to locale()?

the most important arguments are setting what the date_format, time_format, and decimal_mark to what you will then be using for the rest of the data

### Question 2: What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default value of grouping_mark when you set decimal_mark to “,”? What happens to the default value of decimal_mark when you set the grouping_mark to “.”?

```{r 11.3.5 question 2}
locale(decimal_mark = ".", grouping_mark = ".")
```
You get an error message: Error: `decimal_mark` and `grouping_mark` must be different

### Question 3: I didn’t discuss the date_format and time_format options to locale(). What do they do? 

```{r 11.3.5 question 3}
locale()
```
these functions allow you to set the date and time formats manually using the following settings:
Year: %Y (4 digit), %y (2 digit)
Month: %m (2 digit), %b (abbreviated name), %B (full name)
Day: %d (2 digits), %e (optional leading space)
Time: %H (0-23 hr), %I (0-12), %p (AM/PM), %M (minutes), %S (integer seconds), %OS (real seconds), %Z (time zone), %z (offset from UTC, ex +0800)

### Question 4: If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.

```{r 11.3.5 Question 4}
locale(grouping_mark = ".", tz = "GMT")
```

### Question 5: What’s the difference between read_csv() and read_csv2()?

The difference is in the delimiter; read_csv uses a comma, read_csv2 uses a semicolon

### Question 6: What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.

UTF-8 is standard now, before it was ASCII.

For the European languages, there are separate encodings: Latin script, Cyrillic, Greek, Hebrew, Turkish: 
usually with separate ISO and Windows encoding standards. There is also Mac OS Roman.

For Asian languages Arabic and Vietnamese have ISO and Windows standards. The other major Asian scripts have their own:

Japanese: JIS X 0208, Shift JIS, ISO-2022-JP
Chinese: GB 2312, GBK, GB 18030
Korean: KS X 1001, EUC-KR, ISO-2022-KR

### Question 7: Generate the correct format string to parse each of the following dates and times:

```{r 11.3.5 question 7}
d1 <- "January 1, 2010"
parse_date(d1, "%B %d, %Y")

d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%d")

d3 <- "06-Jun-2017"
parse_date(d3, "%d-%b-%Y")

d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d (%Y)")

d5 <- "12/30/14" # Dec 30, 2014
parse_date(d5, "%m/%d/%y")

t1 <- "1705"
parse_time(t1, "%H%M")

t2 <- "11:15:10.12 PM"
parse_time(t2, "%I:%M:%OS %p")
```

