---
title: "R4DS Exercises"
author: "Maddie Lombardo"
date: "1/18/2022"
output: 
    html_document:
      toc: true
      toc_float: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Chapter 3.2.4 Exercises

### Question 1: Run ggplot(data = mpg). What do you see?
```{r 3.2.4 Question 1}
ggplot(data = mpg) 
```

This line of code does not output a graph as there is no geom function to tell R studio what to output. Need '+ geom_point(mapping = aes(x = displ, y = 'variable'))' to see anything

### Question 2: How many rows are in mpg? How many columns?
`r nrow(mpg)` rows, `r ncol(mpg)` columns

### Question 3: What does the drv variable describe? Read the help for ?mpg to find out.
drv = the type of drive train, f = fwd, r = rwd, 4 = 4wd

### Question 4: Make a scatterplot of hwy vs cyl.
```{r 3.2.4 Question 4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

### Question 5: What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r 3.2.4 Question 5}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

This is not helpful because there are no numerical values in class or drv so a scatterplot wouldn't be a good way to visualize this data.

## Chapter 3.3.1 Exercises

### Question 1: What’s gone wrong with this code? Why are the points not blue?
```{r 3.3.1 Question 1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

#this is incorrect because there is a missing parenthesis after hwy. should be:

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

### Question 2:Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

categorical variables: manufacturer, model, trans, drv, fl, class
continuous variables: displ, year, cyl, cty, hwy

you can see this when you run mpg under the columns where it displays the type of variable (cat = <chr> cont = <dbl>, <int>)

### Question 3:Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r 3.3.1 Question 3.1, error = TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = displ, color = year, shape = cyl))
```

you can't map a continuous variable to shape

```{r 3.3.1 Question 3.2}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = manufacturer, y = model, size = model, color = trans, shape = class))
```

You shouldn't use size for a discrete variable; the shape palette can only handle 6 values so any more doesn't work well; asks you to specify shapes if you have to have them.

### Question 4: What happens if you map the same variable to multiple aesthetics?

The points for what you are graphing each have a specific color, size, and shape.
```{r 3.3.1 Question 4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = manufacturer, y = model, size = class, color = class, shape = class))
```

### Question 5: What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

if a shape had a border, stroke would control the width of the border

### Question 6: What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r 3.3.1 Question 6}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = manufacturer, y = model, color = displ < 5))
```

This colors the points based on if the condition displ < 5 is true or false

## Chapter 3.5.1 Exercises

### Question 1: What happens if you facet on a continuous variable?

Creates a new graph for each of the numbers in the continuous variable like displ, see below
```{r 3.5.1 Question 1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ displ, nrow = 2)
```


### Question 2: What happens if you facet on a continuous variable?

The empty cells in the plot with facet_grid(drv ~ cyl) show that there are no points that relate to both variables (cyl, drv). for example, for the cells with 5 cylinders and 4wd, there are no points so cannot plot the displ and hwy for it.
```{r 3.5.1 Question 2}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

### Question 3: What plots does the following code make? What does . do?

The '.' determines where the drv axis will be (on the righthand side vs the top). It is the placeholder for the empty axis.
```{r 3.5.1 Question 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .) 

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ drv) 
```

### Question 4: Take the first faceted plot in this section. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

The advantage of using facet wrap instead of color is that it allows you to more easily see differences in hwy and displ for each car type because they are separated into individual graphs. Especially for large datasets, it would be quite difficult to see all of one car types' data points with color because they could overlap on the graph.
```{r 3.5.1 Question 4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```


```




