
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

# Chapter 4

## Chapter 4.4 Exercises

### Question 1: Why does this code not work?
```{r 4.4 Question 1, error = TRUE}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

The error is that the 'i' in my_variable is a 1 when it is attempted to be called.R needs you to be explicit with no typos when attempting to execute code.

### Question 2: Tweak each of the following R commands so that they run correctly:

```{r 4.4 Question 2, error = TRUE}
ggplot(dota = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) #won't run
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8) #won't run
filter(mpg, cyl == 8)

filter(diamond, carat > 3)#won't run
filter(diamonds, carat > 3)
```

# Chapter 5

## Chapter 5.2.4 Exercises

### Question 1: Find all flights that:

1.1 Had an arrival delay of two or more hours
```{r 5.2.4 Question 1.1}
delay_more_2 <- filter(nycflights13::flights, arr_delay > 120)
```

`r nrow(delay_more_2)` flights had an arrival delay of 2 or more hours

1.2 Flew to Houston (IAH or HOU)
```{r 5.2.4 Question 1.2}
flew_to_houston <- filter(nycflights13::flights, dest == "IAH" | dest == "HOU")
```

`r nrow(flew_to_houston)` flights flew to a Houston airport

1.3 Were operated by United, American, or Delta
```{r 5.2.4 Question 1.3}
un_am_del <- filter(nycflights13::flights, carrier == "AA" | carrier == "DL" | carrier == "UA")
```

`r nrow(un_am_del)` flights were operated by United, American, or Delta

1.4 Departed in summer (July, August, and September)
```{r 5.2.4 Question 1.4}
summer_flight <- filter(nycflights13::flights, month %in% c(7,8,9))
```

`r nrow(summer_flight)` flights departed in summer

1.5 Arrived more than two hours late, but didn’t leave late
```{r 5.2.4 Question 1.5}
arr_2_late_dep_on_time <- filter(nycflights13::flights, dep_delay <= 0 & arr_delay > 120)
```

`r nrow(arr_2_late_dep_on_time)` flights arrived more than 2 hours late, but left on time or early.

1.6 Were delayed by at least an hour, but made up over 30 minutes in flight
```{r 5.2.4 Question 1.6}
made_up_30 <- filter(nycflights13::flights, dep_delay >= 60 & ((dep_delay - arr_delay) > 30))
```

`r nrow(made_up_30)` flights were delayed by at least an hour but made up over 30 minutes in air

1.7 Departed between midnight and 6am (inclusive)
```{r 5.2.4 Question 1.7}
overnight_flight_dep <- filter(nycflights13::flights, dep_time %in% c(12,1,2,3,4,5,6))
```

`r nrow(overnight_flight_dep)` flights left between midnight and 6am

### Question 2: Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

between() will select the rows of values that fall within a specific range. Must be a numeric vector.You could simplify the last exercise (1.7) by:
```{r 5.2.4 Question 2}
overnight_dep <- filter(nycflights13::flights, between(dep_time, 1, 6) | dep_time ==12)
```

### Question 3: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r 5.2.4 Question 3}
missing_dep_time <- filter(nycflights13::flights, is.na(dep_time))
```

`r nrow (missing_dep_time)` flights have a missing departure time

These flights are also missing a dep_delay and arr_time, so these may represent the flights that were cancelled.

### Question 4: Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

NA ^ 0 = 1 because everything to the 0th power is 1.
NA | TRUE it'll still return the the result of the boolean.
FALSE & NA will return the result of the boolean.
The general rule is that it will return the boolean value.
NA*0 = NA because when you try to do math on an NA value, it will return NA

## Chapter 5.3.1 Exercises

### Question 1: How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).

```{r 5.3.1 Question 1}
arrange(flights, desc(is.na(flights)))
```

### Question 2: Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r 5.3.1 Question 2}
most_delay <- arrange(flights, desc(dep_delay))

most_delay %>% 
  arrange(dep_time)
```

### Question 3: Sort flights to find the fastest (highest speed) flights.

```{r 5.3.1 question 3}
flights <- flights %>% 
  mutate(speed = distance/hour)

fastest_flights <- arrange(flights, desc(speed))
```

### Question 4: Which flights travelled the farthest? Which travelled the shortest?

```{r 5.3.1 question 4}
far <- arrange(flights, desc(distance))

short <- arrange(flights, distance)
```

## Chapter 5.4.1 Exercises

### Question 1: Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r 5.4.1 question 1}
flights %>% select(matches("^dep_"),matches("^arr_"))
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)
# you can also select by column position number
flights %>%  select(4,6,7,9)
```
*these are the reasonable ways to do this, you could do ridiculous things like subtracting every name but those you want

### Question 2: What happens if you include the name of a variable multiple times in a select() call?

```{r 5.4.1 question 2}
flights %>% select(year, year, month,day, year)
```

It will only print the variable one time, regardless of how many times you call the variable name in select()

### Question 3: What does the any_of() function do? Why might it be helpful in conjunction with this vector?

```{r 5.4.1 question 3}

vars <- c("year", "month", "day", "dep_delay", "arr_delay")
flights %>%  select(any_of(vars))
```

any_of() select variables in a character vector and does not check for missing variables.

### Question 4: Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r 5.4.1 Question 4}
select(flights, contains("TIME"))
```
No, as the code prints all the variables that contain the string "time" within it. The default is that ignore.case = TRUE, so the capitalization within the code wouldn't effect the output. You can change ignore.case = FALSE to make it case dependent.

## 5.5.2 Exercises

### Question 1: Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r 5.5.2 question 1}
flights <- flights %>% 
  mutate(min_since_mid_dep_time = dep_time %/% 100 * 60 + dep_time %% 100)

flights <- flights %>% 
  mutate(min_since_mid_sched_dep_time = sched_dep_time%/% 100 * 60 + sched_dep_time %% 100)
```

### Question 2: Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r 5.5.2 Question 2.1}
head(flights$air_time)

airtime2 <- flights$arr_time - flights$dep_time
head(airtime2)
```

The air_time's are smaller than the (arr_time-dep_time)'s. This is because the arr_time and dep_time are written not in the minutes since midnight but rather just the hourminutes of time (i.e. 315 = 3:15) together. air_time is the total amount of time spent in the air in minutes. Therefore, to fix this, you should use calculate the minutes