---
title: "chapter_15_16"
author: "Maddie Lombardo"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Chapter 15 Exercises

## Chapter 15.3.1 Exercises

### Question 1: Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r 15.3.1 question 1}
gss_cat <- forcats::gss_cat

gss_cat %>% 
  ggplot(aes(rincome))+
  geom_bar()
```
The labels are too long and overlap so that you can't read them.
Also going to remove the n/a, refused, don't know, no answer responses

```{r 15.3.1 question 1.2}
gss_cat %>% 
  filter(!rincome %in% c("Not applicable", "Don't know", "No answer", "Refused")) %>%
  ggplot(aes(rincome))+
  geom_bar()+
  coord_flip()
```

### Question 2: What is the most common relig in this survey? What’s the most common partyid?

```{r 15.3.1 question 2}
gss_cat %>% 
  count(relig) %>% 
  arrange(-n) %>% 
  head(1)

#most common religion is Protestant

gss_cat %>% 
  count(partyid) %>% 
  arrange(-n) %>% 
  head(1)

#most common party is Independent
```

### Question 3: Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

```{r chapter 15.3.1 question 3}
gss_cat %>% 
    filter(!denom %in% c(
    "No answer", "Other", "Don't know", "Not applicable",
    "No denomination"
  )) %>% 
  count(relig)

gss_cat %>% 
  count(relig, denom) %>% 
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point()
```

## Chapter 15.4.1 Exercises

### Question 1: There are some suspiciously high numbers in tvhours. Is the mean a good summary?

```{r 15.4.1 question 1}
summary(gss_cat["tvhours"])

gss_cat %>% 
  filter(!is.na(tvhours)) %>% 
  ggplot(aes(x = tvhours)) +
  geom_histogram(binwidth = 1)
```

### Question 2: For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.

```{r 15.4.1 question 2}
keep(gss_cat, is.factor) %>% names()

levels(gss_cat$marital)

gss_cat %>%
  ggplot(aes(x = marital)) +
  geom_bar()

levels(gss_cat$race)

gss_cat %>%
  ggplot(aes(race)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE)

levels(gss_cat$rincome)
#this is organized from smallest to largest income but all of the non-answer forms are not together, which is less helpful

levels(gss_cat$partyid)
#ranges from strong republican to strong democrat, showing the spectrum

levels(gss_cat$relig)

gss_cat %>%
  ggplot(aes(relig)) +
  geom_bar() +
  coord_flip()

#no real ordering of religion

levels(gss_cat$denom)
```

### Question 3: Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

```{r 15.4.1 question 3}
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(rincome_summary, aes(age, fct_relevel(rincome, "Not applicable"))) +
  geom_point()
```
It applies a level of 1 to not applicable, putting it at the bottom of the plot

## Chapter 15.5.1 Exercises

### Question 1: How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r 15.5.1 question 1}
gss_cat %>%
  mutate(
    partyid =
      fct_collapse(partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat")
      )
  ) %>%
  count(year, partyid) %>%
  group_by(year) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(
    x = year, y = p,
    colour = fct_reorder2(partyid, year, p)
  )) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID.")
```

### Question 2: How could you collapse rincome into a small set of categories?

```{r 15.5.1 question 2}
gss_cat %>% 
  mutate(rincome = fct_collapse(rincome,
                                na = c("No answer", "Don't know", "Refused", "Not applicable"),
                                below_5000 = c("$4000 to 4999","$3000 to 3999","$1000 to 2999","Lt $1000"),
                                between_5000_15000 = c("$10000 - 14999" ,"$8000 to 9999" ,"$7000 to 7999", "$6000 to 6999" , "$5000 to 5999"),
                                over_15000 = c("$25000 or more", "$20000 - 24999", "$15000 - 19999"))) %>% 
  ggplot(aes(x = rincome))+
  geom_bar()
```

# Chapter 16

## Chapter 16.2.4 Exercises

### Question 1: What happens if you parse a string that contains invalid dates?

```{r 16.2.4 question 1}
library(lubridate)
library(nycflights13)

ymd(c("2010-10-10", "bananas"))
```

You get an NA reply for the invalid data and a warning message of whichever are failing to parse

### Question 2: What does the tzone argument to today() do? Why is it important?

tzone allows you to set which time zone you'd like the current time in. It defaults to the computer system time zone. It is important because if you're attempting to get the date/time you want to get it for the correct location in the world.

### Question 3: Use the appropriate lubridate function to parse each of the following dates:

```{r 16.2.4 question 3}
d1 <- "January 1, 2010"
mdy(d1)

d2 <- "2015-Mar-07"
ymd(d2)

d3 <- "06-Jun-2017"
dmy(d3)

d4 <- c("August 19 (2015)", "July 1 (2015)")
mdy(d4)

d5 <- "12/30/14" # Dec 30, 2014
mdy(d5)
```

## Chapter 16.3.4 Exercises

### Question 1: How does the distribution of flight times within a day change over the course of the year?

```{r 16.3.4 question 1}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt %>%
  filter(!is.na(dep_time)) %>%
  mutate(dep_hour = update(dep_time, yday = 1)) %>%
  mutate(month = factor(month(dep_time))) %>%
  ggplot(aes(dep_hour, color = month)) +
  geom_freqpoly(binwidth = 60 * 60)
```
It looks like the distribution of each month over the course of the day is relatively similar in shape.

### Question 2: Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.

if the dep_delay is accurate, than (dep_time = dep_delay + sched_dep_time)

```{r 16.3.4 question 2}

flights_dt %>%
  mutate(actual_dep_time = sched_dep_time + dep_delay * 60) %>%
  filter(actual_dep_time != dep_time) %>%
  select(actual_dep_time, dep_time, sched_dep_time, dep_delay)
```
if you look at the actual_dep_time vs the dep_time, there are differences in the DATE with which they actually left (they were a day off).

### Question 3: Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)

```{r 16.3.4 question 3}

timezones <- airports %>% 
  select(faa, tzone)

flights_dt %>% 
  inner_join(timezones, by = c(dest = 'faa')) %>% 
  mutate(duration = arr_time-dep_time) %>% 
  mutate(difference = duration - air_time) %>% 
  filter(air_time != duration) %>% 
  filter(difference < 0) %>% 
  select(origin, dest, air_time, duration, difference, tzone) %>% 
  view()

```
For all planes that have a difference of < 0, or saying that there was less time in the air than there was subtracting the arr_time from the dep_time, they are in a difference time zone from departure (NY). This suggests that it doesn't take into consideration a timezone change when calculating the airtime.

### Question 4: How does the average delay time change over the course of a day? Should you use dep_time or sched_dep_time? Why?



### Question 5: On what day of the week should you leave if you want to minimise the chance of a delay?

### Question 6: What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

### Question 7: Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.